---
title: "EDLD654 - Lab 2"
author: "Lea Frank, Stephanie Gluck, Cameron Kay, & Ashley Miller"
date: "10/21/2020"
output:
  html_document:
    theme: spacelab
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)
library(magrittr)

options(scipen = 999)

```

## Read in the `train.csv` data. Please feel free to use `sample_frac()` if you find that the data file is too large for your machine.

```{r, data}
data <- rio::import(here::here("data", "train.csv"))
```


## 1. Initial Split

Set a seed and split the data into a training set and a testing set as two named objects. 

```{r, initial_split}
set.seed(42)

data_split <- initial_split(data)

data_train <- training(data_split)
  
data_test  <- testing(data_split)

```

## 2. Resample

Set a seed and use 10-fold cross-validation to resample the traning data.

```{r, resample}
set.seed(42)

data_train_r10 <- vfold_cv(data_train)

```

## 3. Preprocess

Complete the code maze below by filling in the blanks (____) to create a recipe object that includes:
* a formula model with `score` predicted by 4 predictors
* be sure there are no missing data in your predictors (try `step_naomit()`)
* center and scale all numeric predictors
* dummy code all nominal predictors

```{r, preprocess}

lasso4_rec <- 
  recipe(
    formula = score ~ gndr + ethnic_cd + enrl_grd + econ_dsvntg, 
    data = data_train #use your training set here
  ) %>%
  step_naomit(everything(), skip = TRUE) %>% 
  step_string2factor(gndr, ethnic_cd, econ_dsvntg) %>%  #may not need, depending on your formula
  step_dummy(gndr, ethnic_cd, econ_dsvntg) %>% #may not need, depending on your formula
  step_normalize(enrl_grd) #may not need, depending on your formula

```

## 4. Parsnip model

Create a `{parsnip}` lasso model where the penalty hyperparameter is set to be tuned.

```{r, lasso}

mod_lasso <- linear_reg() %>%
  set_engine("glmnet") %>% 
  set_mode("regression") %>% # redundant; just getting in the habit
  set_args(penalty = tune(), 
           mixture = 1)

```

## 5. Fit a tuned lasso model

Complete the code maze below to fit a tuned lasso model.

```{r, lasso_fit_1}

lasso_grid <- grid_regular(penalty())

lasso4_fit_1 <- tune_grid(
  mod_lasso,
  preprocessor = lasso4_rec,
  resamples = data_train_r10,
  grid = lasso_grid,
  control = tune::control_resamples(verbose = TRUE,
                                    save_pred = TRUE)
)

```

### Question A
  + How many models were fit to each fold of `lasso4_fit_1`? (Please provide a numeric answer, *and* use code to corroborate your answer.)
  
  Since there are 3 penalty values being tested, 3 models will be fitted to each fold.
  
```{r}

lasso4_fit_1 %>%
  collect_metrics(summarize = FALSE) %>%
  filter(id == "Fold01" & .metric == "rmse") %>%
  nrow()

```

  + Use code to list the different values of `penalty()` that were used.

```{r}

lasso_grid

```

## 6. Fit another tuned lasso model

Use your code from (5) above to complete the code maze below to fit a second tuned lasso model, using the same `parsnip` model, `recipe`, and re-sampled object you used before.

```{r, lasso_fit_2}

lasso4_fit_2 <- tune_grid(
  mod_lasso,
  preprocessor = lasso4_rec,
  resamples = data_train_r10,
  control = tune::control_resamples(verbose = TRUE,
                                    save_pred = TRUE)
)

```

### Question B

  + How many models were fit to each fold of `lasso4_fit_2`? (Please provide a numeric answer, *and* use code to corroborate your answer.)
  
    10 

```{r}

lasso4_fit_2 %>%
  collect_metrics(summarize = FALSE) %>%
  filter(id == "Fold01" & .metric == "rmse") %>%
  nrow()

```

  + If this is different than the number of models of `lasso4_fit_1`, please explain why.
  
  Yes,

  + Use code to list the different values of `penalty()` that were used for *lasso4_fit_2*.

```{r}

lasso4_fit_2 %>%
  collect_metrics(summarize = FALSE) %$%
  unique(penalty)

# unique(collect_metrics(lasso4_fit_2, summarize = FALSE)$penalty)

```

## 7. Complete the necessary steps to create and fit a tuned lasso model that has seven or more predictors (use any tuning grid you like). Note that you will need to create a new recipe as well.

```{r, lasso8}

lasso7_rec <- 
  recipe(
    formula = score ~ gndr + ethnic_cd + enrl_grd + econ_dsvntg + lat + lon + dist_sped, 
    data = data_train #training dataset goes here
  ) %>%
  step_naomit(everything(), skip = TRUE) %>% 
  step_string2factor(gndr, ethnic_cd, econ_dsvntg, dist_sped) %>%  #may not need, depending on your formula
  step_dummy(gndr, ethnic_cd, econ_dsvntg, dist_sped) %>% #may not need, depending on your formula
  step_normalize(enrl_grd, lat, lon)

lasso7_fit_2 <- tune_grid(
  mod_lasso,
  preprocessor = lasso7_rec,
  resamples = data_train_r10,
  control = tune::control_resamples(verbose = TRUE,
                                    save_pred = TRUE)
)

```

## 8. Compare the metrics from the best lasso model with 4 predictors to the best lasso model with 7+ predicors. Which is best?

```{r}

lasso4_fit_2 %>%
  show_best(metric = "rmse", n = 1)

lasso7_fit_2 %>%
  show_best(metric = "rmse", n = 1)

```

Our model with 7 predictors is the best fitting model, yielding an rmse of 99.3 (compared to rmse = 100 for the 4 predictor model).

## 9. Fit a tuned elastic net model with the same predictors from (7). 
  + Create a new `{parsnip}` elastic net model
  + Use the same recipe from (7) above
  + Create and apply a regular grid for the elastic net model
  + Compare the metrics from the elastic net model to the best lasso model from (8). Which would you choose for your final model? What are the best hyperparameters for that model?

```{r}

enet_params <- parameters(penalty(), mixture())
enet_grid <- grid_regular(enet_params, levels = c(10, 5)) #want 10 different values of penalty and 5 for mixture
#50 models per fold = 500 models total!

options(scipen = 999) #turns off scientific notation
unique(enet_grid$penalty) #10 unique values for penalty
unique(enet_grid$mixture) #5 unique values for mixture


enet7_tune_rec <- linear_reg() %>%
  set_engine("glmnet") %>% 
  set_mode("regression") %>%
  set_args(penalty = tune(), #placeholder for tuned penalty
           mixture = tune()) #placeholder for tuned mixture

enet7_tune_fit1 <- tune_grid(
  object = enet7_tune_rec,
  preprocessor = lasso7_rec,
  resamples = data_train_r10,
  grid = enet_grid, #grid object specified above
#  metrics = yardstick::metric_set(rmse),
  control = tune::control_resamples(verbose = TRUE,
                                    save_pred = TRUE)
)

lasso7_fit_2 %>%
  show_best(metric = "rmse", n = 1) %$%
  print(mean)

enet7_tune_fit1 %>%
  show_best(metric = "rmse", n = 1) %$%
  print(mean)

```

The elastic net model barely improves model fit (rmse difference is 0.00008). 
